### 2008-10-22 Latex/H2Preview/H2PrevAbstract.tex
My edits. Change to abstract to reflect new objectives and minor corrections to DJNL's changes.

\label{sec:Abstract}
This paper is concerned with a broad class of $\htwo$ controller synthesis problems with preview. A generic preview design problem is 
embedded in a standard generalized regulator framework, and features both previewable and non-previewable disturbances. Preview regulation is acomplished by a two-degree-of-freedom output-feedback controller. The paper is concerned with a number of theoretical issues including the efficient solution of the standard $\htwo$ full-information Riccati equation and the efficient evaluation of the full-information preview gain matrices. The full-information problem is then extended to include the evaluation and efficient synthesis of the output-feedback controller.
 The synthesis of feedforward controllers with preview is analysed as a special case; this problem is of interest to designers who wish to introduce preview as a separate part of a system design. The way in which preview reduces the $\htwo$-norm of the closed-loop system is analyzed in detail.
Closed-loop norm reduction formulae provide a systematic way of establishing how much preview is required to solve a particular problem, and determine when extending the preview horizon will not produce worthwhile benefits. The paper concludes with a summary of the main features of preview control, as well as some controller design insights.

>>>>

\label{sec:Abstract}
This paper is concerned with  providing a generic set of tools for synthesising $\htwo$ preview controllers for a broad range of realistic design problems. A generic preview design problem is 
embedded in a standard generalized regulator framework, and features both previewable and non-previewable disturbances. Preview regulation is acomplished by a two-degree-of-freedom output-feedback controller. The paper is concerned with a number of theoretical issues including the efficient solution of the standard $\htwo$ full-information Riccati equation and the efficient evaluation of the full-information preview gain matrices. The full-information problem is then extended to include the efficient synthesis and efficient implementation of the output-feedback controller.
 The synthesis of feedforward controllers with preview is analysed as a special case; this problem is of interest to designers who wish to introduce preview as a separate part of a system design. The way in which preview reduces the $\htwo$-norm of the closed-loop system is analyzed in detail.
Closed-loop norm reduction formulae provide a systematic way of establishing how much preview is required to solve a particular problem, and determine when extending the preview horizon will not produce worthwhile benefits. The paper concludes with a summary of the main features of preview control, as well as some controller design insights.

### 2008-10-22 Latex/H2Preview/H2PrevIntro.tex
My edits. Change to abstract to reflect new objectives and minor corrections to DJNL's changes.

\label{sec:Intro}
There are many situations in which reference signals, or future disturbances are `previewable'. 
Optimal Preview Control is concerned with designing controllers that exploit previewed information in order to achieve performance levels that are superior to those achievable using current information alone. This paper considers the generic preview synthesis problem illustrated in Figure \ref{fig:DistRejSys}, which comprises a two-degree-of-freedom controller and both previewed disturbances/references ($r$), and unpreviewed disturbances ($w$).
An $\htwo$-optimal solution to this controller synthesis problem is provided that requires only low-dimensional computations and low-dimensional Riccati equation solutions, and leads to a controller whose high-dimensional component is a Finite Impulse Response (FIR) filter; the efficient implementation of FIR filters is well known is the signal processing literature.
The low-dimensional solution to the problem described in Figure~\ref{fig:DistRejSys} derives from the fact that the states of the (high-dimensional) delay line can be reconstructed by making a copy of $\Phi$ in the controller.  The objective of this paper is to provide a framework for synthesizing preview controllers for any problem that fits into the framework illustrated in Figure~\ref{fig:DistRejSys}. In addition, we aim to provide some general insights into the design of preview controllers, and a method for assessing the effectiveness of preview in terms of the achievable $\htwo$-norm reduction.
\begin{figure}
\begin{center}
\stdcontrolfrags
\psfrag{Phi}{$\Phi$}
\psfrag{r}{$r$}
\psfrag{rhat}{$\hat r$}
\psfrag{eta}{$\eta$}
\psfrag{G_p}{$G_p$}
\includegraphics[height=4.0cm]{./diags/DistRejSys.eps}
\caption{\label{fig:DistRejSys} A generalized regulator problem with both previewable and non-previewable disturbances. The transfer function $G$ is the system to be controlled, $K$ is the controller to be synthesized and $\Phi$ is an $N$-step delay line. The disturbance $w$ is not previewable, the control and measurement signals are $u$ and $y$ respectively, $\hat r$ is the previewable disturbance, and $r$ is the future value of $\hat r$. The filter $W_r$ is used  model the expected frequency content of $r$.}
\end{center}
\end{figure}

@@

Figure~\ref{fig:SimpleSISOPrev} illustrates a simple example that may be used to highlight the benefit of preview, the broad structure of the controller, and the effect of preview on the achievable $\htwo$-norm of the closed-loop system. In this formulation, the preview action arises because of the delay line $\Phi$. The input to the controller is $r$, which is the future value of the reference, and $K$ is chosen so as to ensure that $e$ is `small', and hence the plant output follows $\Phi r$ as closely as possible. 
\begin{figure}
\stdcontrolfrags
\begin{center}
\includegraphics[width=0.4\hsize]{./diags/SimpleExampleNominalNoK2.eps}
\end{center}
\caption{A simple SISO open-loop preview tracking problem. The transfer function $\Phi=\z^{-N}$ is an $N$-step delay, $G$ is the plant to be controlled, and $K$ is a feedforward controller. The signal $r$ is the \textit{future} value of the reference, and $e$ is the tracking error.}
\label{fig:SimpleSISOPrev}
\end{figure}
%
Now define the error system:
\als{E(\z)=G(\z)K(\z)-\Phi(\z),}
and assume that $G(\z)$ is stable; in the case that $G(\z)$ is unstable, it could be replaced by $\hat G(\z)=G(\z)(1-G(\z)K_f(\z))^{-1}$ in which $K_f(\z)$ is a stabilising feedback controller. Providing that $G(\z)$ has all its zeros inside the unit circle, perfect tracking ($E(\z)=0$) may be achieved by simply setting $K(\z)=G(\z)^{-1}\Phi(\z)$. 
However, if $G(\z)$ is {NMP}, then such a $K(\z)$ is not internally stabilising and a controller must be found that recognises the limits imposed by {NMP} zeros on the achievable tracking performance. %Such limitations are well known for non-preview systems (i.e. $N=0$ and $\Phi=1$), and are investigated in some detail in \cite{Middleteon_2004_PrevPerf} and \cite{Mirkin_2004_FixedLagPerfSaturation} for the preview case.

For the case where $G(\z)$ is an arbitrary stable, rational transfer function having a single real {NMP} zero at $c_z$, an algebraic expression for the $\htwo$-optimal controller will now be found. Our objective is to find an internally stabilising $K(\z)$ such that $\nrm{E(\z)}_2$ is minimised.

The following inner-outer factorization  may be performed:
\als{G(\z)&= G_o(\z) G_i(\z),
\intertext{where: }
G_i(\z)&=\frac{\z-c_z}{1-\z c_z}
.}
We can write $E(\z)= (\tilde{K}(\z) -\Phi(\z)G_i(\z^{-1}))G_i(\z)$ in which $\tilde{K}(\z)= K(\z) G_o(\z)$ and:
\mbox{$G_i(\z^{-1})G_i(\z)=1$}. The optimal controller is found by setting $K(\z) = (\Phi(\z)G_i(\z^{-1}))_+ G_o^{-1}(\z) $, where $(\cdot)_+$ denotes the stable projection \citep{Doyle_1990_FeedbackControlTheory}. 
It follows by direct calculation  that:
\aln{K(\z)&=\underbrace{G_o(\z)^{-1}}_{IIR}\underbrace{\left(-c_z^{-1}\z^{-N} +  (1-c_z^{-2})c_z \sum^N_{i=1}(\z^{i-N}/c_z^i) \right)}_{FIR},\nonumber
\intertext{and that:}
\nrm{E(\z)}_2&=\frac{1}{\left|c_z^{N+1}\right|}\sqrt{c_z^2-1}.\label{eqn:H2normsimple}}
%

Since $\nrm{E(\z)}_2\rightarrow 0$  as $N\rightarrow \infty$, we conclude that in this example preview action can overcome completely the tracking limitation
imposed by the {NMP} zero. The optimal controller contains a high-order {FIR} part and a low-order {IIR} part, where the preview action comes from the {FIR} part. The dynamics of the {FIR} block is  fully specified by the {RHP} zero, $c_z$, and the preview length, $N$. The fact that the high order part of the controller is FIR leads to an efficient hardware implementation.
%A similar structure is also observed in the solution to the synthesis problem of Figure~\ref{fig:DistRejSys}, and leads to an efficient implementation of the high-order controller.

>>>>

\label{sec:Intro}
There are many situations in which reference signals, or future disturbances are `previewable'. 
Optimal Preview Control is concerned with designing controllers that exploit previewed information in order to achieve performance levels that are superior to those achievable using current information alone. This paper considers the generic preview synthesis problem illustrated in Figure \ref{fig:DistRejSys}, which comprises a two-degree-of-freedom controller and both previewed disturbances/references ($r$), and unpreviewed disturbances ($w$).
An $\htwo$-optimal solution to this controller synthesis problem is provided that requires only low-dimensional computations and low-dimensional Riccati equation solutions, and leads to a controller whose high-dimensional component is a Finite Impulse Response (FIR) filter; the efficient implementation of FIR filters is well known in the signal processing literature.
The low-dimensional solution to the problem described in Figure~\ref{fig:DistRejSys} derives from the fact that the states of the (high-dimensional) delay line can be reconstructed by making a copy of $\Phi$ in the controller.  The objective of this paper is to provide a framework for synthesizing preview controllers for any problem that fits into the framework illustrated in Figure~\ref{fig:DistRejSys}. In addition, we aim to provide some general insights into the design of preview controllers, and a method for assessing the effectiveness of preview in terms of the achievable $\htwo$-norm reduction.
\begin{figure}
\begin{center}
\stdcontrolfrags
\psfrag{Phi}{$\Phi$}
\psfrag{r}{$r$}
\psfrag{rhat}{$\hat r$}
\psfrag{eta}{$\eta$}
\psfrag{G_p}{$G_p$}
\includegraphics[height=4.0cm]{./diags/DistRejSys.eps}
\caption{\label{fig:DistRejSys} A generalized regulator problem with both previewable and non-previewable disturbances. The transfer function $G$ is the system to be controlled, $K$ is the controller to be synthesized and $\Phi = I \z^{-N}$ is an $N$-step delay line (where $\z$ is the Z-transform variable). The disturbance $w$ is not previewable, the control and measurement signals are $u$ and $y$ respectively, $\hat r$ is the previewable disturbance, and $r$ is the future value of $\hat r$. The filter $W_r$ is used  model the expected frequency content of $r$.}
\end{center}
\end{figure}

@@

Figure~\ref{fig:SimpleSISOPrev} illustrates a simple example that may be used to highlight the benefit of preview, the broad structure of the controller, and the effect of preview on the achievable $\htwo$-norm of the closed-loop system. The preview action arises from the delay line $\Phi$. The input to the controller is $r$, which is the future value of the reference, and $K$ is chosen so as to ensure that $e$ is `small', and hence the plant output follows $\Phi r$ as closely as possible. 
\begin{figure}
\stdcontrolfrags
\begin{center}
\includegraphics[width=0.4\hsize]{./diags/SimpleExampleNominalNoK2.eps}
\end{center}
\caption{A simple SISO open-loop preview tracking problem. The transfer function $\Phi=\z^{-N}$ is an $N$-step delay, $G$ is the plant to be controlled, and $K$ is a feedforward controller. The signal $r$ is the \textit{future} value of the reference, and $e$ is the tracking error.}
\label{fig:SimpleSISOPrev}
\end{figure}
%
Define the error system:
\als{E(\z)=G(\z)K(\z)-\Phi(\z),}
and assume that $G(\z)$ is stable; in the case that $G(\z)$ is unstable, it could be replaced by $\hat G(\z)=G(\z)(1-G(\z)K_f(\z))^{-1}$ in which $K_f(\z)$ is a stabilising feedback controller. Providing that $G(\z)$ has all its zeros inside the unit circle, perfect tracking ($E(\z)=0$) may be achieved by simply setting $K(\z)=G(\z)^{-1}\Phi(\z)$. 
However, if $G(\z)$ is nonminimum phase (NMP), then such a $K(\z)$ is not internally stabilising and a controller must be found that recognises the limits imposed by {NMP} zeros on the achievable tracking performance. %Such limitations are well known for non-preview systems (i.e. $N=0$ and $\Phi=1$), and are investigated in some detail in \cite{Middleteon_2004_PrevPerf} and \cite{Mirkin_2004_FixedLagPerfSaturation} for the preview case.

For the case where $G(\z)$ is an arbitrary stable rational transfer function having a single real {NMP} zero at $c_z$, the $\htwo$-optimal controller is easily found. Our objective is to find an internally stabilising $K(\z)$ such that $\nrm{E(\z)}_2$ is minimised.

The following inner-outer factorization  may be performed:
\als{G(\z)&= G_o(\z) G_i(\z),
\intertext{where: }
G_i(\z)&=\frac{\z-c_z}{1-\z c_z}
.}
We can write $E(\z)= (\tilde{K}(\z) -\Phi(\z)G_i(\z^{-1}))G_i(\z)$ in which $\tilde{K}(\z)= K(\z) G_o(\z)$ with
\mbox{$G_i(\z^{-1})G_i(\z)=1$}. The optimal controller is found by setting $K(\z) = (\Phi(\z)G_i(\z^{-1}))_+ G_o^{-1}(\z) $, where $(\cdot)_+$ denotes the stable projection \citep{Doyle_1990_FeedbackControlTheory,LimebeerGreen}. 
It follows by direct calculation  that:
\aln{K(\z)&=\underbrace{G_o(\z)^{-1}}_{IIR}\underbrace{\left(-c_z^{-1}\z^{-N} +  (1-c_z^{-2})c_z \sum^N_{i=1}(\z^{i-N}/c_z^i) \right)}_{FIR},\nonumber
\intertext{and that:}
\nrm{E(\z)}_2&=\frac{1}{\left|c_z^{N+1}\right|}\sqrt{c_z^2-1}.\label{eqn:H2normsimple}}
%

Since $\nrm{E(\z)}_2\rightarrow 0$  as $N\rightarrow \infty$, we conclude that in this example preview action can overcome completely the tracking limitation
imposed by the {NMP} zero. The optimal controller contains a high-order {FIR} part and a low-order infinite impulse response {IIR} part, where the preview action comes from the {FIR} part. The dynamics of the {FIR} block is  fully specified by the {RHP} zero, $c_z$, and the preview length, $N$. The fact that the high order part of the controller is FIR leads to an efficient hardware implementation.
%A similar structure is also observed in the solution to the synthesis problem of Figure~\ref{fig:DistRejSys}, and leads to an efficient implementation of the high-order controller.

### 2008-10-22 Latex/H2Preview/H2PrevNotation.tex
DJNL written conclusions
My edits. Change to abstract to reflect new objectives and minor corrections to DJNL's changes.

\label{sec:notation}
We will make use of discrete-time state-space models of the form:
\als{
x(k+1)&=Ax(k)+Bu(k)\\
y(k)&=Cx(k)+Du(k)
}
in which $k$ is the time index, $x(k)$ is a vector of state variables, $u(k)$ is a vector of inputs, $y(k)$ is a vector of outputs, and $A,B,C$ and $D$ are appropriately dimensioned real matrices. Signals will sometimes be represented by omitting the time index, e.g.
$$x=\{x(k)\}^{\infty}_{-\infty}.$$
When transfer functions are associated with these models, they are computed using:
\als{
G(\mathcal{Z})=C(\mathcal{Z} I -A)^{-1}B+D
}
in which $\mathcal{Z}$ is the Z-transform variable. We will also use the shorthand notation:
\aln{
G(\mathcal{Z})\shorteq\ssmodf{A}{B}{C}{D}\label{eqn:GABCD}.
}
The transfer function $G(\mathcal{Z})$ will be abbreviated $G$ when no confusion will occur.

@@

% The product operator will have the definition:
% \als{\prod_{i=1}^{N} T_i= T_1T_{2}\ldots T_{N-1} T_N }
%The set of eigenvalues of $A$ will be denoted as $\lambda(A)$ and the eigenvalues of the matrix pencil $A-\lambda B$ will be written as $\lambda(A,B)$. 
An $m\times p$ dimensional zero matrix will be denoted as $0_{m\times p}$ and an $n$ dimensional identity matrix will written as $I_n$. The shorthand $0_m=0_{m\times m}$ will also be used. We will often use the notation $\star_{m\times p}$ to represent a matrix belonging to $\mathbb R^{m\times p}$; the dimensions may be omitted when no ambiguity will result.

>>>>

\label{sec:notation}
We will make use of discrete-time state-space models of the form:
\als{
x(k+1)&=Ax(k)+Bu(k)\\
y(k)&=Cx(k)+Du(k)
}
in which $k$ is the time index, $x(k)$ is a vector of state variables, $u(k)$ is a vector of inputs, $y(k)$ is a vector of outputs, and $A,B,C$ and $D$ are appropriately dimensioned real matrices. Signals will sometimes be represented by omitting the time index, e.g.
$$x=\{x(k)\}^{\infty}_{-\infty}.$$
When transfer functions are associated with these models, they are computed using:
\als{
G(\mathcal{Z})=C(\mathcal{Z} I -A)^{-1}B+D
}
in which $\mathcal{Z}$ is the Z-transform variable. We will also use the shorthand notation:
\aln{
G(\mathcal{Z})\shorteq\ssmodf{A}{B}{C}{D}\label{eqn:GABCD}.
}
The transfer function $G(\mathcal{Z})$ will be abbreviated by $G$ when no confusion will occur.

@@

% The product operator will have the definition:
% \als{\prod_{i=1}^{N} T_i= T_1T_{2}\ldots T_{N-1} T_N }
%The set of eigenvalues of $A$ will be denoted as $\lambda(A)$ and the eigenvalues of the matrix pencil $A-\lambda B$ will be written as $\lambda(A,B)$. 
An $m\times p$ dimensional zero matrix will be denoted as $0_{m\times p}$ and an $n$ dimensional identity matrix will written as $I_n$. The shorthand $0_m=0_{m\times m}$ will also be used. %We will often use the notation $\star_{m\times p}$ to represent a matrix belonging to $\mathbb R^{m\times p}$; the dimensions may be omitted when no ambiguity will result.
